Journey: A Study in Unhelpful Titles (TODO: should we get a new title?)

### 2017-04-27 paper.tex
Moved related work
Updated methodology section

@@

We conduct a series of experiments to measure the scalability of various parts
of the Linux memory management system.  First, we measure the scalability of
Linux's \texttt{vm\_area\_struct}. Then, we measure the minimum amount of memory
necessary for swapping.

\begin{figure}
\centering
\begin{tabular}{|l|p{5cm}|} \hline
OS & Ubuntu 16.04.2 \\ \hline
Kernel & Linux 4.4.0-70 \\ \hline
CPU & Intel Xeon E5645, 2.40 GHz, 6 cores/12 threads, 384KB Private L1-I Cache,
1.536 Unified L2 Cache, 12MB Shared L3 Cache \\ \hline
Memory & 16GB, 1066MHz \\ \hline
Storage & 500GB, 7200RPM \\
\hline
\end{tabular}
\caption{Specifications of our test machine.  \label{fig:specs}}
\end{figure}

The specifications of the machine we used to run our experiments are listed in
Figure \ref{fig:specs}. We disabled Intel SpeedStep and TurboBoost and ensure
that the processor frequency stays constant (TODO: did we do this?). All
experiments are pinned to the same core on our machine.

We run several different microbenchmarks, each of which has a different memory
allocation pattern. These benchmarks are designed to stress the memory
management system in different ways, rather than simulate real workloads. 
Each run is executed in isolation, without any additional processes
running beyond services that start when the system is booted and a single ssh
session. We are careful to avoid opening extra ssh sessions, running screen or
tmux, or executing extra commands while a benchmark is running, as these may
impact memory usage, and thus our results.

We measured latency using the \texttt{rdtsc} instruction provided by x86\_64
processors, which gives high resolution cycle-level timestamps. We take a
timestamp before and after each memory management operation in the workload and
take the difference to get latency. (TODO: is this still true?)

To measure memory usage, we use the procfs's reporting on current memory usage.
Where appropriate, we record the slab allocator's count of active
\texttt{vm\_area\_struct}s in the system. Measuring kernel memory usage and
associating memory usage with processes is difficult. Our methodology assumes
that our benchmarks are running in an otherwise idle system and should dominate
increases in memory usage.  We compute total memory usage by the rest of the
system as the total amount of memory less the amount of memory used by the
benchmark.

Finally, we keep transparent huge pages on throughout our experiments, unless
otherwise specified.

@@

Assume the memory usage of the rest of the system is $R$. We denote $R_i$ as the
$i$-th measurement of $R$ during a run ($i = 0, 1, ... 2^{20}$). System memory
usage is known to jitter in Linux (TODO: cite). To adjust for this jitter, we
define $\Delta_i = R_i - R_0$. For each operation, we take the median value of
$\Delta_i$ across the 5 runs to reduce jitter further.  This is the value
graphed in the figures in the following section. $\Delta_i$ represents the
increase in memory usage due to memory management overheads caused by running
$i$ memory management operations.

@@

\item \textit{Fragmented} (``frag'' for short). This workload first allocates
    $4^{20}$ contiguous 4KB blocks. Then, it resizes every other block to 16KB.
        The resizing operation, rather than the initial allocation is measured.
\end{itemize}

@@

We run this experiment with both a strided access pattern and a random access
pattern. The stride access pattern simply touches pages in order of increasing
address. The random access pattern chooses a random address. Both patterns touch
the same number of addresses.

After each page of virtual memory is touched, the resident set size (RSS; the
amount of physical memory Linux has allocated to the process) is recorded.  The
values of RSS are stored in an mlocked array and written to a file after the
benchmark completes. This prevents the writing of the results from interacting
with paging. These RSS values are graphed in our results section.

@@

A single process allocates 18GB of virtual memory space. It then proceeds to
write one byte of each 4KB page in the region, leading to the actual allocation
of physical memory. The process chooses addresses randomly, as in the previous
set of experiments, inducing the maximum amount of swapping.

In all these experiments, the system is given 32GB of swap space total. We
measure the impact of physical memory size and transparent huge pages. We run
four experiments:

\begin{enumerate}
    \item 1GB physical memory with huge pages enabled
    \item 16GB physical memory with huge pages enabled
    \item 1GB physical memory with huge pages disabled
    \item 16GB physical memory with huge pages disabled
\end{enumerate}

To run the 1GB memory benchmarks, we hot unplug 15GB using the sysfs in Linux.
This makes the unplugged memory unusable by the whole system (kernel and user).

We measure CPU time via the procfs using the same methodology as the
\texttt{top} utility. The \texttt{/proc/[pid]/stat} file for kswapd is read.
From this file the \texttt{utime} and \texttt{stime} fields are read and summed
to get the total time allocated to kswapd. Then, the \texttt{/proc/stat} file is
read to get the total system time elapsed. By dividing, we can find the portion
of CPU time spent in kswapd.

\section{Result}

@@

>>>>

(TODO: swapping large pages?)

@@

We conduct a series of experiments to measure the scalability of various parts
of the Linux memory management system.  First, we measure the scalability of
Linux's \texttt{vm\_area\_struct}. Then, we measure the scalability of Linux's
swapping mechanisms.

\begin{figure}
\centering
\begin{tabular}{|l|p{5cm}|} \hline
OS & Ubuntu 16.04.2 \\ \hline
Kernel & Linux 4.4.0-70 \\ \hline
CPU & Intel Xeon E5645, 2.40 GHz, 6 cores/12 threads, 32kB L1-D, 32kB L1-I,
    256kB L2, 12MB L3 \\ \hline
Memory & 16GB, 1066MHz \\ \hline
Storage & 500GB HDD, 7200RPM \\
\hline
\end{tabular}
\caption{Specifications of our test machine.  \label{fig:specs}}
\end{figure}

The specifications of the machine we used to run our experiments are listed in
Figure \ref{fig:specs}. We disabled Intel SpeedStep and TurboBoost and ensure
that the processor frequency stays constant. All experiments are pinned to the
same core on our machine.

We run several different microbenchmarks, each of which has a different memory
allocation pattern. These benchmarks are designed to stress the memory
management system in different ways, rather than simulate real workloads. 
Each run is executed in isolation, without any additional processes
running beyond services that start when the system is booted and a single ssh
session. We are careful to avoid opening extra ssh sessions, running screen or
tmux, or executing extra commands while a benchmark is running since these may
impact memory usage and, thus, our results.

We measured latency using the \texttt{rdtsc} instruction provided by x86\_64
processors, which gives high resolution cycle-level timestamps. We take a
timestamp before and after each memory management operation in the workload and
take the difference to get latency.

To measure memory usage, we use the procfs's reporting on current memory usage.
Where appropriate, we record the slab allocator's count of active
\texttt{vm\_area\_struct}s in the system. Measuring kernel memory usage and
associating memory usage with processes is difficult. Our methodology assumes
that our benchmarks are running in an otherwise idle system and should dominate
increases in memory usage.  We compute total memory usage by the rest of the
system as the total amount of memory less the amount of memory used by the
benchmark. We keep transparent huge pages on throughout our experiments, unless
otherwise specified.

To measure processor usage, we use the procfs's reporting on number of jiffies.
\texttt{/proc/stat} reports the number of jiffies of uptime the system has seen
overall, while \texttt{/proc/[pid]/stat} reports on the specified process.

@@

Assume the memory usage of the rest of the system is $R$. We denote $R_i$ as the
$i$-th measurement of $R$ during a run ($i = 0, 1, ..., 2^{20}$). System memory
usage is known to jitter in Linux (TODO: cite). To adjust for this jitter, we
define $\Delta_i = R_i - R_0$. For each operation, we take the median value of
$\Delta_i$ across the 5 runs to reduce jitter further.  This is the value
graphed in the figures in the following section. $\Delta_i$ represents the
increase in memory usage due to memory management overheads caused by running
$i$ memory management operations.

@@

\item \textit{Fragmented} (``frag'' for short). This workload first allocates
    $2^{21}$ contiguous 4KB blocks. Then, it resizes every other block to 16KB.
        The resizing operation, rather than the initial allocation is measured.
\end{itemize}

@@

We run this experiment with both a continuous access pattern and a random access
pattern. The continuous access pattern simply touches adjacent pages in order of
increasing address. The random access pattern chooses a random address. Both
patterns touch the same number of addresses.

After each page of virtual memory is touched, the resident set size (RSS; the
amount of physical memory Linux has allocated to the process) is recorded.  The
values of RSS are stored in an mlocked array and written to a file after the
benchmark completes. This prevents the writing of the results from interacting
with paging in the kernel page cache. These RSS values are graphed in our
results section.

@@

A single process allocates 18GB of virtual memory space and writes one byte of
each 4KB page in the region. The same two access patterns are tested as in the
previous set of experiments (continuous and random).

In all these experiments, the system is given 32GB of swap space total. We
repeat each experiment with hugepages enabled and disabled to see the impact of
the transparent hugepage system. The processor usage of kswapd and khugepaged is
recorded for each experiment and reported in the next section.

\section{Results}

@@

TODO: khugepaged?
